Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-06 ms.tex
Added subsection describing how we predict wp from point data, resolving Issue #7.

We model the radial velocity distribution of satellite galaxies as a Gaussian with 
first moment equal to the host halo velocity and second moment equal to the 
solution to the isotropic Jeans equation for an NFW profile \citep{more09b}, 
\begin{equation}
\sigma^{2}_{r}(\tilde{r}|c) = V_{\rm vir}^{2}\frac{c^{2}\tilde{r}(1 + c\tilde{r})^{2}}{g(c)}\int_{c\tilde{r}}^{\infty}{\rm d}y\frac{g(y)}{y^{3}(1 + y)^{2}}, 
\end{equation}
where $\tilde{r} = r/\rvir,$ $g(x) = \rm{ln}(1+x) - x / (1+x),$ and $V_{\rm vir}^{2} = G\mvir/\rvir.$ 
We assume that velocities are isotropic, setting peculiar velocities in the simulation $z$-dimension  
equal to random draws from the above distribution. We place mock galaxies 
into redshift-space by replacing each galaxy's $z$-coordinate with $z \rightarrow z + V_{\rm z}/H_0.$

@@

We have re-analyzed the SDSS DR7 measurements of projected galaxy clustering,  $\wprp,$ and number density, $n_{\rm g},$ originally published in \citet{zehavi_etal11}. Our work is especially novel in that we provide the first quantitative constraints on assembly bias derived from the Decorated HOD, an extension to the traditional HOD introduced in \citet{hearin_etal16} developed for exactly this purpose. 
We enumerate our most important conclusions below. 
\ben
\item It is not possible to rule out galaxy assembly bias using SDSS DR7 measurements of $\wprp$ and $n_{\rm g}.$
\item Decorated HOD fits to $\wprp$ and $n_{\rm g}$ favor significant levels of assembly bias, particularly in the lower luminosity thresholds we study. Both the $\magr<-20, -20.5$ samples prefer relatively strong central galaxy assembly bias, while at lower luminosities, the $\magr<-19$ sample favors satellite assembly bias. 
\item Galaxy assembly bias generally weakens for brighter galaxy samples: Decorated HOD fits to the $\magr<-21$ sample are consistent with both $\abias^{\rm cens}=0$ and $\abias^{\rm sats}=0.$ This consistent with the predictions made in \citet{hearin_etal16} on purely theoretical grounds that galaxy assembly bias should generically weaken with increasing galaxy mass. 
\item Our posteriors and best-fit parameters summarized in Table \ref{table:parameters} supersede the values published in \citet{zehavi_etal11}, as direct-mock-population together with the Decorated HOD allows us to account for highly significant systematics that have heretofore been neglected from all HOD fits to SDSS data. We note that our findings update the original \citet{zehavi_etal11} results {\em even for our fits in which assembly bias has been fixed to zero}, since the original results derive from unconverged MCMC chains that do not sufficiently sample the HOD model posteriors. 
\een
We conclude by noting that since $\wprp$ and $n_{\rm g}$ are already very well-measured in DR7, it is likely that further improvements on assembly bias constraints at low-redshift will require additional observational measurements, e.g., galaxy--galaxy lensing, group statistics, etc.

>>>>

We model the radial velocity distribution of satellite galaxies as a Gaussian with 
first moment equal to the host halo velocity and second moment equal to the 
solution to the isotropic Jeans equation for an NFW profile \citep{more09b}, 
\begin{equation}
\sigma^{2}_{r}(\tilde{r}|c) = V_{\rm vir}^{2}\frac{c^{2}\tilde{r}(1 + c\tilde{r})^{2}}{g(c)}\int_{c\tilde{r}}^{\infty}{\rm d}y\frac{g(y)}{y^{3}(1 + y)^{2}}, 
\end{equation}
where $\tilde{r} = r/\rvir,$ $g(x) = \rm{ln}(1+x) - x / (1+x),$ and $V_{\rm vir}^{2} = G\mvir/\rvir.$ 
We assume that velocities are isotropic, setting the peculiar velocities in each Cartesian direction 
according to random draws from the above radial velocity distribution.

\subsubsection{Predictions for observables}

After populating a halo catalog with mock galaxies, we calculate the comoving number density of our mock galaxy sample as $n_{\rm g} = N_{\rm gal} / L_{\rm box}^3.$ We apply the distant-observer approximation and use the simulation z-axis as the line-of-sight direction, and the distance between points in the $xy$-plane to define the projected distance $\rp.$ We place mock galaxies into redshift-space by replacing each galaxy's $z$-coordinate with $z \rightarrow z + V_{\rm z}/H_0.$ We count pairs of points in each of our $\rp$ bins, rejecting pairs with $z-$distance exceeding $\pi_{\rm max}=60$ ${\rm Mpc/h.}$ We employ the \citet{landyszalay93} estimator to turn pair counts into a $\wprp$ prediction, as described in detail in {\tt halotools.readthedocs.io}.

@@

We have re-analyzed the SDSS DR7 measurements of projected galaxy clustering,  $\wprp,$ and number density, $n_{\rm g},$ originally published in \citet{zehavi_etal11}. Our work is especially novel in that we provide the first quantitative constraints on assembly bias derived from the Decorated HOD, an extension to the traditional HOD introduced in \citet{hearin_etal16} developed for exactly this purpose. 
We enumerate our most important conclusions below. 
\ben
\item It is not possible to rule out galaxy assembly bias using SDSS DR7 measurements of $\wprp$ and $n_{\rm g}.$
\item Decorated HOD fits to $\wprp$ and $n_{\rm g}$ favor significant levels of assembly bias, particularly in the lower luminosity thresholds we study. Both the $\magr<-20, -20.5$ samples prefer relatively strong central galaxy assembly bias, while at lower luminosities, the $\magr<-19$ sample favors satellite assembly bias. 
\item Galaxy assembly bias generally weakens for brighter galaxy samples: Decorated HOD fits to the $\magr<-21$ sample are consistent with both $\abias^{\rm cens}=0$ and $\abias^{\rm sats}=0.$ This consistent with the well-established result that {\em halo assembly bias} weakens with increasing halo mass over the dynamic range relevant to these galaxy samples \citep[see, e.g., Figure 8 of][and references therein]{hearin_etal16}. 
\item Our posteriors and best-fit parameters summarized in Table \ref{table:parameters} supersede the values published in \citet{zehavi_etal11}, as direct-mock-population together with the Decorated HOD allows us to account for highly significant systematics that have heretofore been neglected from all HOD fits to SDSS data. We note that our findings update the original \citet{zehavi_etal11} results {\em even for our fits in which assembly bias has been fixed to zero}, since the original results derive from unconverged MCMC chains that do not sufficiently sample the HOD model posteriors. 
\een
We conclude by noting that since $\wprp$ and $n_{\rm g}$ are already very well-measured in DR7, it is likely that further improvements on assembly bias constraints at low-redshift will require additional observational measurements, e.g., galaxy--galaxy lensing, group statistics, etc.
